Preregistration Experiment 6: Is convergence more trustworthy when there are more choice options?

### 2023-05-22 Experiment_4/analysis.Rmd
add experiment 6 and slight model updates

```{r}
# model for accuracy
# random intercept and slope by participants
model_accuracy <- lmer(accuracy ~ convergence + (1 + convergence | id), 
                                 data = d)

summary(model_accuracy)
```

@@

```{r}
# model for competence
# random intercept and slope by participants
model_competence <- lmer(competence ~ convergence + 
                           (1 + convergence | id), data = d)

summary(model_competence)
```

@@

In the models above, we treated convergence as a continuous variable. Based on the different levels, we will build a categorical variable, `convergence_categorical`.

```{r}
# make a categorical variable from `convergence`
d <- d %>% 
  mutate(convergence_categorical = recode_factor(convergence, 
                                                 `0` = "opposing majority", 
                                                 `1` = "divergence", 
                                                 `2` = "majority", 
                                                 `3` = "consensus",
                                                 .default = NA_character_)
         )

levels(d$convergence_categorical)

```

We run the same models outlined in the hypotheses section, but replacing `convergence` with `convergence_categorical`. This also allows us to inspect heterogeniety in differences between levels (with respect to the baseline, i.e. "opposing majority").

>>>>

```{r}
# model for accuracy
# random intercept and slope by participants
model_accuracy <- lmer(accuracy ~ convergence + (1 + convergence | id), 
                                 data = d)
```

@@

```{r}
# model for competence
# random intercept and slope by participants
model_competence <- lmer(competence ~ convergence + 
                           (1 + convergence | id), data = d)
```

```{r}
# combined results table
modelsummary(list("Accuracy" = model_accuracy , "Competence" = model_competence), 
             title = 'Models estimating effect of Convergence', 
             stars = TRUE)
```

@@

In the models above, we treated convergence as a continuous variable. We run the same models outlined in the hypotheses section, but replacing `convergence` with `convergence_categorical` (where all levels are categories). This also allows us to inspect heterogeniety in differences between levels (with respect to the baseline, i.e. "opposing majority").

```{r}
# show levels of `convergence`
levels(as.factor(d$convergence_categorical))
```

```{r}
# model for accuracy
# random intercept and slope by participants
model_accuracy <- lmer(accuracy ~ convergence_categorical + (1 + convergence_categorical | id), 
                                 data = d)
# model for competence
# random intercept and slope by participants
model_competence <- lmer(competence ~ convergence_categorical + 
                           (1 + convergence_categorical | id), data = d)
```

```{r}
# combined results table
modelsummary(list("Accuracy" = model_accuracy , "Competence" = model_competence), 
             title = 'Models estimating effect of Convergence', 
             stars = TRUE)
```

### 2023-05-22 Experiment_4/cleaning.Rmd
add experiment 6 and slight model updates

## Re-shape data
```{r}
d <- d %>% 
  # add an easy to read participant identifier
  mutate(id = 1:nrow(.)) %>% 
  # bring to long format
  pivot_longer(cols = c("0_acc_a_1" : "3_com_b"), 
               names_to = "condition", values_to = "score") %>% 
  # For some weird reason, qualtrics added `_1` at the end of each accuracy
  # stimulus name. We remove that
  mutate(across(condition, ~str_remove(., "_1"))) %>% 
  # separate conditions into CONVERGENCE_OUTCOME_STIMULUS
  separate_wider_delim(condition, "_", names = c("convergence", "outcome", 
                                                 "stimulus_variant")
                       ) %>%
  pivot_wider(names_from = outcome, values_from = score) %>% 
  # create better variable names
  rename(competence = com, 
         accuracy = acc) %>% 
  # all variables are coded as `character` - make key variables numeric
  mutate(across(c(convergence, competence, accuracy), as.numeric)
         )
```

>>>>

## Re-shape data
```{r}
d <- d %>% 
  # add an easy to read participant identifier
  mutate(id = 1:nrow(.)) %>% 
  # bring to long format
  pivot_longer(cols = c("0_acc_a_1" : "3_com_b"), 
               names_to = "condition", values_to = "score") %>% 
  # For some weird reason, qualtrics added `_1` at the end of each accuracy
  # stimulus name. We remove that
  mutate(across(condition, ~str_remove(., "_1"))) %>% 
  # separate conditions into CONVERGENCE_OUTCOME_STIMULUS
  separate_wider_delim(condition, "_", names = c("convergence", "outcome", 
                                                 "stimulus_variant")
                       ) %>%
  pivot_wider(names_from = outcome, values_from = score) %>% 
  # create better variable names
  rename(competence = com, 
         accuracy = acc) %>% 
  # all variables are coded as `character` - make key variables numeric
  mutate(across(c(convergence, competence, accuracy), as.numeric), 
         # make a categorical variable from `convergence`
         convergence_categorical = recode_factor(convergence, 
                                                 `0` = "opposing majority", 
                                                 `1` = "divergence", 
                                                 `2` = "majority", 
                                                 `3` = "consensus",
                                                 .default = NA_character_)
         )
```

### 2023-05-22 Experiment_4/model.Rmd
add experiment 6 and slight model updates

We define competence simply as the probability of making an accurate choice (among various options). We assume that individuals (i) vary in competence,(ii) that competence is continuous and ranges from pure chance (`1/options`) to certainly making the correct choice (`1`), and that it is distributed uniformly in the population.

```{r}
# Imagine a situation with 3 choice options
options <- 3
# generate a uniform competence distribution for a population of n = 10'000
population <- 10000
competence = runif(population, min = 1/options, max = 1)

# data generating function
ggplot() +
  stat_function(fun = dunif, args = list(min = 0, max = 1)) +
  labs(title = "Uniform distribution of competence from 1/3 to 1", x = "Competence", y = "P(Competence)") + 
  xlim(1/3, 1)

# generated population
ggplot() +
  geom_histogram(aes(x = competence)) + 
  scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.1)) + 
  annotate("text", x = 0.9, y = 100, label = paste0("n = ", population), color = "red") +
  labs(title = "(sampled) Population of competence drawn from \n uniform population distribution", 
       y = "Count")
```
### Generate choices from competence

@@

  # 2: randomly draw individual answers based on competence levels 
  data$answer <- data$competence %>% 
    purrr::map_chr(function(x){

@@

```{r}
data_3_sample %>% group_by(sample, constellation) %>% summarise(across(c(average_accuracy, average_relative_competence), mean))

>>>>

We define competence as the probability of making an accurate choice (among various options). Competence is continuous and ranges from `0` to `1`, where `0` corresponds to pure chance (`1/options`) and `1` corresponds to certainly making the correct choice.

In our code this measure is called `relative_competence`, while `competence` designates the actual probability of making an accurate choice, ranging from `1/options` to `1`). For example, in a 3-choice-options scenario, an individual with a probability of 1/3 (`competence == 0.333`) to pick the right answer has `relative_competence == 0`.

We assume that individuals (i) vary in competence and (ii) that competence that it is distributed uniformly in the population.

```{r}
# Imagine a situation with 3 choice options
options <- 3
# generate a uniform competence distribution for a population of n = 10'000
population <- 10000
data <- data.frame(competence = runif(population, min = 1/options, max = 1))

# data generating function
ggplot() +
  stat_function(fun = dunif, args = list(min = 0, max = 1)) +
  labs(title = "Uniform distribution of competence from 0 to 1", x = "Competence", y = "P(Competence)") + 
  xlim(0, 1)

# generated population
ggplot(data, aes(x = relative_competence)) +
  geom_histogram() + 
  scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.1)) + 
  annotate("text", x = 0.9, y = 100, label = paste0("n = ", population), color = "red") +
  labs(title = "(sampled) Population of competence drawn from \n uniform population distribution", 
       y = "Count")
```
### Generate choices from competence

@@

  # 2: Draw individual answers based on competence levels 
  data$answer <- data$competence %>% 
    purrr::map_chr(function(x){

@@

```{r}
regression_data <- data_3_options %>% 
  filter(options %in% c(3, 10)) %>% 
  group_by(iteration, constellation, options) %>% 
  summarise(across(c(average_accuracy, average_relative_competence), mean)) %>% 
  mutate(
    convergence = case_when( 
      constellation == "minority" ~ 0, 
      constellation == "dissensus" ~ 1,
      constellation == "majority" ~ 2,
      constellation == "consensus" ~ 3), 
    options = as.factor(options)
  )

accuracy <- lm(average_accuracy ~ convergence + options, 
               data = regression_data) #%>% broom.mixed::tidy()
competence <- lm(average_relative_competence ~ convergence + options, data = regression_data) %>% broom.mixed::tidy()

summary(accuracy)
```

```{r}
data_3_options %>% group_by(sample, constellation) %>% summarise(across(c(average_accuracy, average_relative_competence), mean))

### 2023-05-22 Experiment_6/analysis.Rmd
add experiment 6 and slight model updates

---
title: 
- "Analysis Experiment 5: Is convergence less trustworthy when informants are biased?"
author: 
- "Hugo Mercier, Jan Pfänder" 
date: "2023-02-02"
output: 
  html_document: 
    keep_md: yes
---

>>>>

---
title: 
- "Analysis Experiment 6: Is convergence less trustworthy when informants are biased?"
author: 
- "Hugo Mercier, Jan Pfänder" 
date: "2023-02-02"
output: 
  html_document: 
    keep_md: yes
---
